---
title: "NIH Aging Data Exploration"
author: "Ben A. Flanagan"
date: "1/10/2018: updated 2/28/2018: updated Sys.Date()"
output: html_document
editor_options: 
  chunk_output_type: console
---
### Notes:

There are a few issues with these data. How are we going to deal with the sampling at the various time points? I could imagine subtracting the number sampled from the total at day 28 and all subsequent counts to develop proportions. That may be the most conservative approach, but I am still unsure. I am still fishing for ideas as how to deal with the sampling artifact. 

In the following analyses, I used ALL data; not just crosses deemed 'usable'. 

```{r setup, include=FALSE}
library(dplyr); library(reshape2); library(ggplot2); library(survival); library(ranger); library(survival); library(ggfortify); library(survminer); library(ggsignif)
rm(list = ls())
std <- function(x) sd(x)/sqrt(length(x))


fe_dat <- read.csv("~/Documents/phd/NIH/longevity/data/NIH_hatching_baf_sheet.csv")
dat<-read.csv("/Users/benjaminflanagan/Documents/phd/NIH/NIH_data_final_melt.csv") # likely errors propgotaing though data as 0s
dat$pops <- substr(dat$cross,start = 0,stop = 2)
dat$day <- as.numeric(substr(dat$day,start = 2,stop = 10))+28


datmast <- 
  dat %>% 
    group_by(day, cross, clutch, sex) %>%
      filter(!any(num_alive=="NA")) ### filters within each day, cross, clutch and sex to find NAs and removes

datmast_use <- subset(datmast, usability=="usable")
```



```{r usability check, echo=FALSE}

dat2<-subset(datmast,day==28 & clutch == 1 & sex == "m") ### to generate one row for each cross

print(table(dat2$pops))
print(table(dat2$usability, dat2$pops)) # current summary 

```


## Sex ratio 

For the following plots I removed any specific sex within a clutch at a given day where we had no data.  

Below are sex specific counts and sex ratio wrapped over days. 




```{r sex ratio all, echo=FALSE, message=FALSE, warning=FALSE}
d28 <- subset(datmast, day==28)

ggplot(d28, aes(x=pops, y=num_alive, fill=sex)) + geom_boxplot() +
    #facet_grid(~Group, scales="free", space="free_x") +
  geom_signif(vjust=0.7,
              textsize=10,
              size=1)

ggboxplot(d28, x="sex", y="num_alive", fill="sex", facet.by = "pops") +
  stat_compare_means(aes(group = sex), label = "p.format") +
  theme_linedraw() +
  ylab("Count per Clutch")+
  xlab("Sex")+
  guides(color=guide_legend(title=c("Sex"))) +
  theme(text=element_text(face="bold", size=20),
        strip.background = element_rect(fill="white"))+
  theme(strip.text = element_text(colour = 'black')) 


ggplot(datmast, aes(x=pops,y=as.numeric(num_alive), fill=sex))+
  geom_boxplot()+ 
  theme_linedraw() +
  facet_wrap(~day)+
  guides(fill=guide_legend(title=c("Sex")))+
  theme(text=element_text(face="bold", size=20))+
  ylab("Count per Clutch")+
  xlab("Cross")+
  ggtitle("All")

tot1 <-
  datmast %>% group_by(sex,pops,day)  %>%
  summarise(total=sum(num_alive))

tot2 <-
  datmast %>% group_by(day,pops) %>%
  summarise(total2=sum(num_alive))

propmast<-full_join(tot1,tot2)
propmast$sex_ratio<-propmast$total/propmast$total2

ggplot(subset(propmast, day==28), aes(x=pops, y=sex_ratio, fill=sex)) +
  geom_col() + 
  theme_linedraw() +
  ylab("Proportion")+
  xlab("Cross")+
  guides(fill=guide_legend(title=c("Sex")))+
  theme(text=element_text(face="bold", size=20))+
  ylim(c(0,1))+
  theme(text=element_text(face="bold", size=20))+
  ggtitle("All")


 
ggplot(subset(propmast), aes(x=pops, y=sex_ratio, fill=sex)) + 
  geom_col()+
  theme_linedraw() +
  facet_wrap(~day)+
  ylab("Proportion")+
  xlab("Cross")+
  guides(fill=guide_legend(title=c("Sex")))+
  theme(text=element_text(face="bold", size=20))+
  theme(text=element_text(face="bold", size=20))+
  ylim(c(0,1))+
  ggtitle("All")

```


```{r sex ratio usable, echo=F,  message=FALSE, warning=FALSE, eval=FALSE}

ggplot(subset(datmast_use, day==28), aes(x=pops,y=as.numeric(num_alive), color=sex))+
  geom_boxplot()+
  ylab("Count per Clutch")+
  xlab("Cross")+
  theme(text=element_text(face="bold", size=20)) +
  ggtitle("Usable")+
  guides(color=guide_legend(title=c("Sex")))+
  theme(text=element_text(face="bold", size=20))

ggplot(datmast_use, aes(x=pops,y=as.numeric(num_alive), color=sex))+
  geom_boxplot()+
  facet_wrap(~day)+
  ylab("Count per Clutch")+
  xlab("Cross")+
  ggtitle("Usable")



tot1_use <-
  datmast_use %>% group_by(day,pops,sex)  %>%
  summarise(total=sum(num_alive))


tot2_use <-
  datmast_use %>% group_by(day,pops) %>%
  summarise(total2=sum(num_alive))

propmast_use<-full_join(tot1_use,tot2_use)
propmast_use$sex_ratio<-propmast_use$total/propmast_use$total2

ggplot(subset(propmast_use, day==28), aes(x=pops, y=sex_ratio, color=sex, fill=sex)) + 
  geom_col()+
  ylab("Proportion")+
  xlab("Cross")+
  ylim(c(0,1))+
  theme(text=element_text(face="bold", size=20))+
  ggtitle("Usable")

ggplot(subset(propmast_use), aes(x=pops, y=sex_ratio, color=sex, fill=sex)) + 
  geom_col()+
  facet_wrap(~day)+
  ylab("Proportion")+
  xlab("Cross")+
  ylim(c(0,1))+
  ggtitle("Usable")

```


```{r heterotic productivyt, echo=FALSE, warning=FALSE}

ggplot(subset(datmast, day==28), aes(x=pops,y=as.numeric(num_alive)))+
  geom_boxplot() + 
  theme_linedraw() +
  ylab("Count per Clutch")+
  xlab("Cross")+
  guides(fill=guide_legend(title=c("Sex")))+
  theme(text=element_text(face="bold", size=20))+
  ggtitle("All")

tmp <- subset(datmast, day == 28)

tmp$parental <- ifelse(tmp$pops == "FF" , "Parental", "Hybrid")
tmp$parental <- ifelse(tmp$pops == "SS" , "Parental", tmp$parental)
tmp$num_alive <- as.numeric(tmp$num_alive)

ggplot(tmp, aes(x = parental, y = num_alive, fill = parental, group = parental)) +
  geom_boxplot(fill=c("#E69F00", "#56B4E9"), color="black") +
  geom_signif(comparisons = list(c("Parental","Hybrid")), map_signif_level = FALSE, position = "identity", textsize = 7) + 
  theme_linedraw() +
  ylab("Count per Clutch") +
  xlab("Cross") +
  guides(fill=FALSE) +
  ylim(c(0,80))+
  theme(text=element_text(face="bold", size=20))

t.test(subset(tmp, parental=="Parental")$num_alive, subset(tmp, parental=="Hybrid")$num_alive)

```



## Survival plots

The following are using all raw data. Again missing data for number alive was removed. All proprotions relevant to total alive at day 28. I have marked the weeks where samples were taken (49-56 & 105-112).

```{r survival plots, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
tot1 <-
  dat %>% group_by(day,pops,sex)  %>%
  summarise(total=sum(na.omit(num_alive)))

tot2 <-
  dat %>% group_by(day,pops) %>%
  summarise(total2=sum(na.omit(num_alive)))

tot3<-tot1[tot1$day==28,]
colnames(tot3)[4]<-"total_d28"

propmast_long_sex<-merge(tot3,tot1, by = c("sex", "pops"))
propmast_long_sex$day.x <- NULL
colnames(propmast_long_sex)[4]<-"day"
propmast_long_sex$prop_alive<-propmast_long_sex$total/propmast_long_sex$total_d28 ### data frame by sex
propmast_long_sex[, ncol(propmast_long_sex)][propmast_long_sex[, ncol(propmast_long_sex)] == 0] <- NA

tot4<-tot2[tot2$day==28,]
colnames(tot4)[3]<-"total_d28"

propmast_long<-merge(tot4,tot2, by = "pops")
propmast_long$day.x <- NULL
colnames(propmast_long)[3]<-"day"
propmast_long$prop_alive<-propmast_long$total2/propmast_long$total_d28### data frame by sex
propmast_long[, ncol(propmast_long)][propmast_long[, ncol(propmast_long)] == 0] <- NA


ggplot(propmast_long, aes(x=day,y=prop_alive,color=pops))+
  geom_vline(xintercept = c(49,56, 105, 112), linetype="dashed", alpha=0.5)+
  geom_point(size=2)+
  geom_line()+
  ylab("Proportion Alive")+
  xlab("Day")+
  ylim(c(0,1))+
  guides(color=guide_legend(title=c("cross")))+
  theme(text=element_text(face="bold", size=20))+
  ggtitle("All")


ggplot(propmast_long_sex, aes(x=day,y=prop_alive,color=sex))+
  #geom_vline(xintercept = c(49,56, 105, 112), linetype="dashed", alpha=0.5)+
  geom_point(size=2)+
  geom_line()+
  facet_wrap(~pops)+
  ylab("Proportion Alive")+
  xlab("Day")+
  ylim(c(0,1))+
  guides(color=guide_legend(title=c("sex")))+
  theme(text=element_text(face="bold", size=15))+
  ggtitle("All")

```


```{r survival plots with usable, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
tot1 <-
    dat %>% filter(usability=="usable") %>%
      group_by(day,pops,sex)  %>%
        summarise(total=sum(na.omit(num_alive)))

tot2 <-
  dat %>% filter(usability=="usable") %>%
    group_by(day,pops) %>%
      summarise(total2=sum(na.omit(num_alive)))

tot3<-tot1[tot1$day==28,]
colnames(tot3)[4]<-"total_d28"

propmast_long_sex<-merge(tot3,tot1, by = c("sex", "pops"))
propmast_long_sex$day.x <- NULL
colnames(propmast_long_sex)[4]<-"day"
propmast_long_sex$prop_alive<-propmast_long_sex$total/propmast_long_sex$total_d28 ### data frame by sex
propmast_long_sex[, ncol(propmast_long_sex)][propmast_long_sex[, ncol(propmast_long_sex)] == 0] <- NA

tot4<-tot2[tot2$day==28,]
colnames(tot4)[3]<-"total_d28"

propmast_long<-merge(tot4,tot2, by = "pops")
propmast_long$day.x <- NULL
colnames(propmast_long)[3]<-"day"
propmast_long$prop_alive<-propmast_long$total2/propmast_long$total_d28### data frame by sex
propmast_long[, ncol(propmast_long)][propmast_long[, ncol(propmast_long)] == 0] <- NA


ggplot(propmast_long, aes(x=day,y=prop_alive,color=pops))+
  geom_vline(xintercept = c(49,56, 105, 112), linetype="dashed", alpha=0.5)+
  geom_point(size=2)+
  geom_line()+
  ylab("Proportion Alive")+
  xlab("Day")+
  ylim(c(0,1))+
  guides(color=guide_legend(title=c("cross")))+
  theme(text=element_text(face="bold", size=20))+
  ggtitle("Usable")


ggplot(propmast_long_sex, aes(x=day,y=prop_alive,color=sex))+
  geom_vline(xintercept = c(49,56, 105, 112), linetype="dashed", alpha=0.5)+
  geom_point(size=2)+
  geom_line()+
  facet_wrap(~pops)+
  ylab("Proportion Alive")+
  xlab("Day")+
  #ylim(c(0,1))+
  guides(color=guide_legend(title=c("sex")))+
  theme(text=element_text(face="bold", size=15))+
  ggtitle("Usable")

```


```{r survival plots with variation due to family, echo=FALSE, message=FALSE, warning=FALSE}
tot1 <-
  dat %>% 
  group_by(day,sex,cross)  %>%
  summarise(total=sum(na.omit(num_alive)))

tot3<-tot1[tot1$day==28,]
colnames(tot3)[4]<-"total_d28"

propmast_long_sex<-merge(tot3,tot1, by = c("sex", "cross")) %>% arrange(sex, cross, day.y)
propmast_long_sex$day.x <- NULL
colnames(propmast_long_sex)[4]<-"day"
propmast_long_sex$prop_alive<-propmast_long_sex$total/propmast_long_sex$total_d28 ### data frame by sex
propmast_long_sex[, ncol(propmast_long_sex)][propmast_long_sex[, ncol(propmast_long_sex)] == 0] <- 0

propmast_long_sex$pops <- substr(propmast_long_sex$cross,0,2)
tmp<- na.omit(propmast_long_sex)

propmast_clutch <-
  tmp %>% group_by(pops, day) %>% summarise(mean_prop=mean(prop_alive), std=std(prop_alive)) 

ggplot(subset(propmast_clutch), aes(x=day,y=mean_prop,color=pops))+
  geom_errorbar(aes(ymin=mean_prop-std, ymax=mean_prop+std), width=.1)+
  #geom_vline(xintercept = c(49,56, 105, 112), linetype="dashed", alpha=0.5)+
  geom_point(size=2)+
  theme_linedraw() +
  geom_line() +
  ylab("Proportion Alive")+
  xlab("Day")+
  ylim(c(0.0000001,1))+
  guides(color=guide_legend(title=c("Cross")))+
  theme(text=element_text(face="bold", size=20))+
  ggtitle("All")

propmast_clutch_sex <-
  tmp %>% group_by(pops, day, sex) %>% summarise(mean_prop=mean(prop_alive), std=std(prop_alive)) 

ggplot(subset(propmast_clutch_sex), aes(x=day,y=mean_prop,color=sex))+
  geom_errorbar(aes(ymin=mean_prop-std, ymax=mean_prop+std), width=.1)+
  #geom_vline(xintercept = c(49,56, 105, 112), linetype="dashed", alpha=0.5)+
  theme_linedraw() +
  geom_point(size=2) +
  geom_line() +
  ylab("Proportion Alive") +
  xlab("Day") +
  ylim(c(0.0000001,1)) +
  guides(color=guide_legend(title=c("Sex"))) +
  theme(text=element_text(face="bold", size=20),
        strip.background =element_rect(fill="white"))+
  theme(strip.text = element_text(colour = 'black')) +
  facet_wrap(~pops) +
  ggtitle("All")


```


```{r survival plots with variation due to family with usable, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
tot1 <-
  dat %>% filter(usability=="usable") %>%
     group_by(day,sex,cross)  %>%
      summarise(total=sum(na.omit(num_alive)))

tot3<-tot1[tot1$day==28,]
colnames(tot3)[4]<-"total_d28"

propmast_long_sex<-merge(tot3,tot1, by = c("sex", "cross"))
propmast_long_sex$day.x <- NULL
colnames(propmast_long_sex)[4]<-"day"
propmast_long_sex$prop_alive<-propmast_long_sex$total/propmast_long_sex$total_d28 ### data frame by sex
propmast_long_sex[, ncol(propmast_long_sex)][propmast_long_sex[, ncol(propmast_long_sex)] == 0] <- 0

propmast_long_sex$pops <- substr(propmast_long_sex$cross,0,2)
tmp<- na.omit(propmast_long_sex)

propmast_clutch <-
  tmp %>% group_by(pops, day) %>% summarise(mean_prop=mean(prop_alive), std=std(prop_alive)) 

ggplot(subset(propmast_clutch), aes(x=day,y=mean_prop,color=pops))+
  geom_errorbar(aes(ymin=mean_prop-std, ymax=mean_prop+std), width=.1)+
  geom_vline(xintercept = c(49,56, 105, 112), linetype="dashed", alpha=0.5)+
  geom_point(size=2)+
  geom_line()+
  ylab("Proportion Alive")+
  xlab("Day")+
  ylim(c(0.0000001,1))+
  guides(color=guide_legend(title=c("cross")))+
  theme(text=element_text(face="bold", size=20))+
  ggtitle("Usable")

propmast_clutch_sex <-
  tmp %>% group_by(pops, day, sex) %>% summarise(mean_prop=mean(prop_alive), std=std(prop_alive)) 

ggplot(subset(propmast_clutch_sex), aes(x=day,y=mean_prop,color=sex))+
  geom_errorbar(aes(ymin=mean_prop-std, ymax=mean_prop+std), width=.1)+
  #geom_vline(xintercept = c(49,56, 105, 112), linetype="dashed", alpha=0.5)+
  theme_linedraw() +
  geom_point(size=2) +
  geom_line() +
  ylab("Proportion Alive") +
  xlab("Day") +
  ylim(c(0.0000001,1)) +
  guides(color=guide_legend(title=c("Sex"))) +
  theme(text=element_text(face="bold", size=20),
        strip.background =element_rect(fill="white"))+
  theme(strip.text = element_text(colour = 'black')) +
  facet_wrap(~pops) +
  ggtitle("Usable")


```


```{r how about breaking the data into two part for sampling periods with all, echo=F, message=FALSE, warning=FALSE, include=FALSE, eval=FALSE}
dat_28 <- subset(dat, day<49)

tot1 <-
  dat_28 %>% 
     group_by(day,sex,cross)  %>%
      summarise(total=sum(na.omit(num_alive))) %>%
        filter(total!=0)

tot3<-tot1[tot1$day==28,]
colnames(tot3)[4]<-"total_d28"

propmast_long_sex<-merge(tot3,tot1, by = c("sex", "cross"))
propmast_long_sex$day.x <- NULL
colnames(propmast_long_sex)[4]<-"day"
propmast_long_sex$prop_alive<-propmast_long_sex$total/propmast_long_sex$total_d28 ### data frame by sex
propmast_long_sex[, ncol(propmast_long_sex)][propmast_long_sex[, ncol(propmast_long_sex)] == 0] <- NA

propmast_long_sex$pops <- substr(propmast_long_sex$cross,0,2)
tmp<- na.omit(propmast_long_sex)


propmast_clutch <-
  tmp %>% group_by(pops, day) %>% summarise(mean_prop=mean(prop_alive), std=std(prop_alive)) 

ggplot(propmast_clutch, aes(x=day,y=mean_prop,color=pops))+
  geom_errorbar(aes(ymin=mean_prop-std, ymax=mean_prop+std), width=.1)+
  geom_point(size=2)+
  geom_line()+
  ylab("Proportion Alive")+
  xlab("Day")+
  ylim(c(0,1))+
  guides(color=guide_legend(title=c("cross")))+
  theme(text=element_text(face="bold", size=20))+
  ggtitle("All")

propmast_clutch_sex <-
  tmp %>% group_by(pops, day, sex) %>% summarise(mean_prop=mean(prop_alive), std=std(prop_alive)) 

ggplot(propmast_clutch_sex, aes(x=day,y=mean_prop,color=sex))+
  geom_errorbar(aes(ymin=mean_prop-std, ymax=mean_prop+std), width=.1)+
  geom_point(size=2)+
  geom_line()+
  ylab("Proportion Alive")+
  xlab("Day")+
  guides(color=guide_legend(title=c("cross")))+
  theme(text=element_text(face="bold", size=20))+
  facet_wrap(~pops)+
  ggtitle("All")


dat_56 <- subset(dat, day>49 & day<104)

tot1 <-
  dat_56 %>%
     group_by(day,sex,cross)  %>%
      summarise(total=sum(na.omit(num_alive))) %>%
        filter(total!=0)

tot3<-tot1[tot1$day==56,]
colnames(tot3)[4]<-"total_d28"

propmast_long_sex<-merge(tot3,tot1, by = c("sex", "cross"))
propmast_long_sex$day.x <- NULL
colnames(propmast_long_sex)[4]<-"day"
propmast_long_sex$prop_alive<-propmast_long_sex$total/propmast_long_sex$total_d28 ### data frame by sex
propmast_long_sex[, ncol(propmast_long_sex)][propmast_long_sex[, ncol(propmast_long_sex)] == 0] <- NA

propmast_long_sex$pops <- substr(propmast_long_sex$cross,0,2)
tmp<- na.omit(propmast_long_sex)


propmast_clutch <-
  tmp %>% group_by(pops, day) %>% summarise(mean_prop=mean(prop_alive), std=std(prop_alive)) 

ggplot(propmast_clutch, aes(x=day,y=mean_prop,color=pops))+
  geom_errorbar(aes(ymin=mean_prop-std, ymax=mean_prop+std), width=.1)+
  geom_point(size=2)+
  geom_line()+
  ylab("Proportion Alive")+
  xlab("Day")+
  guides(color=guide_legend(title=c("cross")))+
  theme(text=element_text(face="bold", size=20))+
  ggtitle("All")

propmast_clutch_sex <-
  tmp %>% group_by(pops, day, sex) %>% summarise(mean_prop=mean(prop_alive), std=std(prop_alive)) 

ggplot(propmast_clutch_sex, aes(x=day,y=mean_prop,color=sex))+
  geom_errorbar(aes(ymin=mean_prop-std, ymax=mean_prop+std), width=.1)+
  geom_point(size=2)+
  geom_line()+
  ylab("Proportion Alive")+
  xlab("Day")+
  ylim(c(0,1))+
  guides(color=guide_legend(title=c("cross")))+
  theme(text=element_text(face="bold", size=20))+
  facet_wrap(~pops)+
  ggtitle("All")


```


```{r how about breaking the data into two part for sampling periods with usable, echo=F, message=FALSE, warning=FALSE, include=FALSE, eval=FALSE}
dat_28 <- subset(dat, day<49)

tot1 <-
  dat_28 %>% filter(usability=="usable") %>%
     group_by(day,sex,cross)  %>%
      summarise(total=sum(na.omit(num_alive))) %>%
        filter(total!=0)

tot3<-tot1[tot1$day==28,]
colnames(tot3)[4]<-"total_d28"

propmast_long_sex<-merge(tot3,tot1, by = c("sex", "cross"))
propmast_long_sex$day.x <- NULL
colnames(propmast_long_sex)[4]<-"day"
propmast_long_sex$prop_alive<-propmast_long_sex$total/propmast_long_sex$total_d28 ### data frame by sex
propmast_long_sex[, ncol(propmast_long_sex)][propmast_long_sex[, ncol(propmast_long_sex)] == 0] <- NA

propmast_long_sex$pops <- substr(propmast_long_sex$cross,0,2)
tmp<- na.omit(propmast_long_sex)


propmast_clutch <-
  tmp %>% group_by(pops, day) %>% summarise(mean_prop=mean(prop_alive), std=std(prop_alive)) 

ggplot(propmast_clutch, aes(x=day,y=mean_prop,color=pops))+
  geom_errorbar(aes(ymin=mean_prop-std, ymax=mean_prop+std), width=.1)+
  geom_point(size=2)+
  geom_line()+
  ylab("Proportion Alive")+
  xlab("Day")+
  ylim(c(0,1))+
  guides(color=guide_legend(title=c("cross")))+
  theme(text=element_text(face="bold", size=20))+
  ggtitle("Usable")

propmast_clutch_sex <-
  tmp %>% group_by(pops, day, sex) %>% summarise(mean_prop=mean(prop_alive), std=std(prop_alive)) 

ggplot(propmast_clutch_sex, aes(x=day,y=mean_prop,color=sex))+
  geom_errorbar(aes(ymin=mean_prop-std, ymax=mean_prop+std), width=.1)+
  geom_point(size=2)+
  geom_line()+
  ylab("Proportion Alive")+
  xlab("Day")+
  ylim(c(0,1))+
  guides(color=guide_legend(title=c("cross")))+
  theme(text=element_text(face="bold", size=20))+
  facet_wrap(~pops)+
  ggtitle("Usable")


dat_56 <- subset(dat, day>49 & day<104)

tot1 <-
  dat_56 %>% filter(usability=="usable") %>%
     group_by(day,sex,cross)  %>%
      summarise(total=sum(na.omit(num_alive))) %>%
        filter(total!=0)

tot3<-tot1[tot1$day==56,]
colnames(tot3)[4]<-"total_d28"

propmast_long_sex<-merge(tot3,tot1, by = c("sex", "cross"))
propmast_long_sex$day.x <- NULL
colnames(propmast_long_sex)[4]<-"day"
propmast_long_sex$prop_alive<-propmast_long_sex$total/propmast_long_sex$total_d28 ### data frame by sex
propmast_long_sex[, ncol(propmast_long_sex)][propmast_long_sex[, ncol(propmast_long_sex)] == 0] <- NA

propmast_long_sex$pops <- substr(propmast_long_sex$cross,0,2)
tmp<- na.omit(propmast_long_sex)


propmast_clutch <-
  tmp %>% group_by(pops, day) %>% summarise(mean_prop=mean(prop_alive), std=std(prop_alive)) 

ggplot(propmast_clutch, aes(x=day,y=mean_prop,color=pops))+
  geom_errorbar(aes(ymin=mean_prop-std, ymax=mean_prop+std), width=.1)+
  geom_point(size=2)+
  geom_line()+
  ylab("Proportion Alive")+
  xlab("Day")+
  ylim(c(0,1))+
  guides(color=guide_legend(title=c("cross")))+
  theme(text=element_text(face="bold", size=20))+
  ggtitle("Usable")

propmast_clutch_sex <-
  tmp %>% group_by(pops, day, sex) %>% summarise(mean_prop=mean(prop_alive), std=std(prop_alive)) 

ggplot(propmast_clutch_sex, aes(x=day,y=mean_prop,color=sex))+
  geom_errorbar(aes(ymin=mean_prop-std, ymax=mean_prop+std), width=.1)+
  geom_point(size=2)+
  geom_line()+
  ylab("Proportion Alive")+
  xlab("Day")+
  ylim(c(0,1))+
  guides(color=guide_legend(title=c("cross")))+
  theme(text=element_text(face="bold", size=20))+
  facet_wrap(~pops)+
  ggtitle("Usable")


```


```{r sex specific max longevity, echo=F}

long <- datmast_use %>% group_by(cross, clutch, sex) %>% filter(num_alive==0) %>% summarise(max_longevity= min(day))
long$pops <- substr(long$cross,start = 0,stop = 2)
long$parental <- ifelse(long$pops == "FF" , "Parental", "Hybrid")
long$parental <- ifelse(long$pops == "SS" , "Parental", long$parental)

ggplot(long, aes(x = pops, y = max_longevity, fill = sex)) +
  geom_boxplot() +
  theme_linedraw() +
  ylab("Max Longevity") +
  xlab("Cross") +
  guides(color=guide_legend(title=c("Sex"))) +
  theme(text=element_text(face="bold", size=20))

ggplot(long, aes(x = parental, y = max_longevity, fill = parental)) +
  geom_boxplot(fill=c("#E69F00", "#56B4E9"), color="black") +  
  geom_signif(comparisons = list(c("Parental","Hybrid")), map_signif_level = FALSE, position = "identity", textsize = 5) + 
  theme_linedraw() +
  ylab("Max Longevity") +
  xlab("Cross") +
  ylim(c(0,280))+
  guides(fill=FALSE) +
  theme(text=element_text(face="bold", size=20))

 ggplot(long, aes(x = parental, y = max_longevity, fill = sex)) +
  geom_boxplot(color="black") +
  theme_linedraw() +
  ylab("Max Longevity") +  
  xlab("Cross") +
  guides(fill=FALSE, color = guide_legend(title=c("Sex"))) +
  theme(text=element_text(face="bold", size=20))

summary(aov(max_longevity~sex*pops, data = long))
TukeyHSD(aov(max_longevity~sex*pops, data = long))

sf_max <- subset(long, parental=="SF")$max_longevity
fs_max <- subset(long, parental=="FS")$max_longevity
p_max <- subset(long, parental=="Parental")$max_longevity

fit <- aov(max_longevity ~ parental * sex, data=long)
summary(fit)
TukeyHSD(fit)
```



```{r coorelation of max and fe}
colnames(fe_dat)[1] <- "cross"
cl1 <- abs(fe_dat$cl1_hatch_day - fe_dat$cl1_fe_day)
cl2 <- abs(fe_dat$cl2_hatch_day - fe_dat$cl2_fe_day) 
cl3 <- abs(fe_dat$cl3_hatch_day - fe_dat$cl3_fe_day)

dfa <- data.frame(dev = c(cl1,cl2, cl3),
                  clutch = c(rep(1,length(cl1)), rep(2, length(cl2)), rep(3, length(cl3))),
                  cross = rep(fe_dat$cross, 3)
                  )

long$max_longevity <- as.integer(long$max_longevity)

tmp1 <- full_join(long, dfa, by = c("cross", "clutch"))
datmast2 <- full_join(datmast, tmp1, by = c("cross", "clutch"))
datmast2$pops <- substr(datmast2$cross,1,2)

ggplot(na.omit(datmast2), aes(x = dev, y = max_longevity)) +
  geom_smooth(method = "lm") +
  geom_point(aes(color = pops)) +
  theme_linedraw() +
  xlab("Development Time") +
  ylab("Max Longevity") +
  theme(text=element_text(face="bold", size=20)) +
  guides(color=guide_legend(title=c("Cross"))) 



fit <- lm(max_longevity~dev , na.omit(datmast2))
summary(fit)

cor.test(datmast2$max_longevity, datmast2$dev, method=c("pearson"))
```


```{r how many died from hatch to splitting correlation with sex ratio, include=T, echo=F}
colnames(fe_dat)[1] <- "cross"
cl1 <- fe_dat$cl1_num
cl2 <- fe_dat$cl2_num
cl3 <- fe_dat$cl3_num

dfa <- data.frame(num_hatched = c(cl1,cl2, cl3),
                  clutch = c(rep(1,length(cl1)), rep(2, length(cl2)), rep(3, length(cl3))),
                  cross = rep(fe_dat$cross, 3)
                  )

datmast_sexsum <- datmast_use %>% filter(day == 28) %>% group_by(clutch, cross) %>%
  summarise(tot_num_alive = sum(num_alive), prop_f = (num_alive[2] / sum(num_alive)))

tmp1 <- full_join(datmast_sexsum, dfa, by = c("cross", "clutch"))
tmp1$pops <- substr(tmp1$cross, 1, 2)
tmp1$diff = with(tmp1, num_hatched-tot_num_alive)
tmp1 <- filter(tmp1, diff >= 0 )

ggplot(tmp1, aes(x=diff, y = prop_f)) +
  geom_point(aes(color = pops)) +
  #geom_smooth(method = "lm") +
  theme_linedraw() +
  xlab("Death in First 28 Days") +
  ylab("Family Sex Ratio") +
  theme(text=element_text(face="bold", size=20)) +
  guides(color=guide_legend(title=c("Cross"))) 

summary(lm(diff~prop_f, data = subset(tmp1, pops == "SS")))

```



## Statistics CPH & K-M

```{r coxph and K-M, include=T, echo=F}
tmp <- NULL
tmp1 <- NULL
tmp2 <- NULL
tmp3 <- NULL
tmp4 <- NULL
for (i in unique(datmast_use$cross)){ ### creates a data frame with ecah indivudal in a single row
  clt <- datmast_use %>% filter(cross==i) %>% group_by(day, sex) %>% summarise(num_alive=sum(num_alive))
    for (k in 1:2){
      soi <- unique(datmast_use$sex)[k]
      clt_sex <- clt %>% filter(sex==soi)  %>% arrange(day)
        for (j in 1:nrow(clt_sex)){
          if(j==1) next
          day<-clt_sex$day[j]
          tot_alive <- clt_sex$num_alive[1]
          if(tot_alive==0){
            print("breaking")
            break
            }else
          diff <- clt_sex$num_alive[j - 1] - clt_sex$num_alive[j]
          if(diff==0) next
          vec <- c(rep(TRUE, diff))
          tmp <- data.frame(rep(i,diff), rep(1:diff), rep(day, diff), rep(soi, diff))
          colnames(tmp)<- c("cross","individual","Ri","sex")
          tmp$Li <- tmp$Ri - 7
          tmp2<-rbind(tmp2,tmp)
      }
      tmp3<-rbind(tmp3,tmp2)
    }
  tmp4<-rbind(tmp4,tmp3)
}

tmp4$pops <- substr(tmp4$cross,start = 0,stop = 2)
summary(as.numeric(subset(tmp4, pops=="FF")$Ri))


km2 <- with(tmp4, Surv(Li, Ri, type = "interval2")) # interval with (,]. I think this is correct

tmp11 <- survfit(km2 ~ pops, data=tmp4)
summary(tmp11)

tmp12 <- survfit(km2 ~ pops + sex, data=tmp4)
summary(tmp12)

ggsurvplot(tmp11, data = tmp4,  main = "Survival curve",
           conf.int = TRUE,
    font.main = c(20, "bold", "black"),
    font.x = c(20, "bold", "black"),
    font.y = c(20, "bold", "black"),
    font.tickslab = c(12, "plain", "black"),
    ggtheme = theme_linedraw(),
    legend = "none") +
  xlab("Day")

ggsurvplot(tmp12, data = tmp4,  main = "Survival curve",
           conf.int = TRUE,
    font.main = c(20, "bold", "black"),
    font.x = c(20, "bold", "black"),
    font.y = c(20, "bold", "black"),
    font.tickslab = c(12, "plain", "black"),
    facet.by = "pops") +
  theme_linedraw() +
  xlab("Day") +
  guides(color=guide_legend(title=c("Sex"))) +
  theme(text=element_text(face="bold", size=20),
        strip.background =element_rect(fill="white"))+
  theme(strip.text = element_text(colour = 'black')) +
  theme(legend.position="none") +
  facet_wrap(~pops) # maybe an error here becuase FS females don't start at 0. It is becasuse we have interval censored data. If an animial died within the week it assumes it died mid-week. 


  

```

The surviorsip plots here using the km model and the porportion alive above likely don't agree becasue the first shot took the mean from each family and the variation was due to family. :)